---
output: github_document
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-",
  cache = T
)
```

# TCIApathfinder

TCIApathfinder is a wrapper for The Cancer Imaging Archive's REST API v3.
The Cancer Imaging Archive (TCIA) hosts de-identified medical images
of cancer available for public download, as well as rich metadata
for each image series. TCIA provides a REST API for programmatic access
to the data. This package provides simple functions to access each API 
endpoint. For more information about TCIA, see TCIA's [website](http://www.cancerimagingarchive.net/).

TCIApathfinder is on [CRAN](https://CRAN.R-project.org/package=TCIApathfinder).

## Installation

From within R:

```{r eval = F}
install.packages("TCIApathfinder")
```

From GitHub:

```{r eval = F}
# install.packages("devtools")
devtools::install_github("pamelarussell/TCIApathfinder")
```

## Authentication

An API key is required to access data from TCIA. To obtain and correctly store your API key:

1. Request a key from TCIA by following the instructions [here](https://wiki.cancerimagingarchive.net/display/Public/TCIA+Programmatic+Interface+%28REST+API%29+Usage+Guide).

2. Create a text file in your home directory (`~/`) called `.Renviron`. 

3. Create the contents of the `.Renviron` file like this, making sure the last line in the file is empty. Otherwise, R will silently fail to load the file.
  
    ```
    TCIA_API_KEY=xxx-xxx-xxx-xxx
    
    ```

4. Restart R. `.Renviron` is only processed at the beginning of an R session.

## Package usage

Detailed vignettes on package usage and downstream image analysis can be viewed on [CRAN](https://CRAN.R-project.org/package=TCIApathfinder) or from within an R session with `browseVignettes("TCIApathfinder")`.

## Citation

TBA. Check out our [biorXiv preprint](https://www.biorxiv.org/content/early/2017/12/30/240986).

## More information on the TCIA REST API

- [API usage guide](https://wiki.cancerimagingarchive.net/display/Public/TCIA+Programmatic+Interface+%28REST+API%29+Usage+Guide)
- [Object type definitions](https://wiki.cancerimagingarchive.net/display/Public/TCIA+API+Return+Values)





